---
title: "Assignment_2"
author: '"Pallavi"'
date: "2023-09-14"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Summary

## Problem Statement

Universal bank is a young bank growing rapidly in terms of overall customer acquisition.
The majority of these customers are liability customers (depositors) with varying sizes of relationship with the bank. The customer base of asset customers (borrowers) is quite
small, and the bank is interested in expanding this base rapidly in more loan business. In particular, it wants to explore ways of converting its liability customers to personal loan customers.

A campaign that the bank ran last year for liability customers showed a healthy conversion rate of over 9% success. This has encouraged the retail marketing department to devise smarter campaigns with better target marketing. The goal is to use k-NN to predict whether a new customer will accept a loan offer. This will serve as the basis for the design of a new campaign.

The file UniversalBank.csv contains data on 5000 customers. The data include customer
demographic information (age, income, etc.), the customer’s relationship with the bank
(mortgage, securities account, etc.), and the customer response to the last personal loan
campaign (Personal Loan). Among these 5000 customers, only 480 (= 9.6%) accepted the
personal loan that was offered to them in the earlier campaign.

Partition the data into training (60%) and validation (40%) sets

## Questions - Answers

1. Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 =
1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1, and
Credit Card = 1. Perform a k-NN classification with all predictors except ID and ZIP code
using k = 1. Remember to transform categorical predictors with more than two categories
into dummy variables first. Specify the success class as 1 (loan acceptance), and use the
default cutoff value of 0.5. How would this customer be classified?

## Load required Libraries
```{r}
library(class)
library(caret)
library(e1071)
```

# Data Cleaning

```{r}
universal.df <-read.csv("C:/Users/palla/OneDrive/Desktop/Assignments/FML/Asignment 2/UniversalBank.csv")
```


```{r}
dim(universal.df)
```


```{r}
head(universal.df)
```

Drop Variable ID and ZIP
```{r}
universal.df<- universal.df[,-c(1,5)]
```

#education need to be converted to factor
```{r}
universal.df$Education<-as.factor(universal.df$Education)
```

#Convert education to dummy variables
```{r}
groups <- dummyVars(~., data=universal.df)
```


```{r}
universal_m.df <- as.data.frame(predict(groups,universal.df))
```

#split the data in to training (60%) and validation(40%)
```{r}
set.seed(1)
```

```{r}
train.index<-sample(row.names(universal_m.df),0.6*dim(universal_m.df)[1])
train.index
```


```{r}
valid.index<-setdiff(row.names(universal_m.df),train.index)
```


```{r}
train.df<-universal_m.df[train.index,]
valid.df<-universal_m.df[valid.index,]
t(t(names(train.df)))
```
#normalising the data (Standaridastion the data )(note that personal income is the 10th variable
```{r}
train.norm.df <- train.df[, -10]
valid.norm.df <- valid.df[, -10]
norm.values <-preProcess(train.df[, -10],method=c("center","scale"))
train.norm.df<-predict(norm.values,train.df[,-10])
valid.norm.df<-predict(norm.values,valid.df[,-10])

```
#we have converted all catergorical variables to dummy variables
#lets create a new sample
```{r}
new_customer<- data.frame(
Age=40,
Experience=10,
Income=84,
Family=2,
CCAvg=2,
Education.1=0,
Education.2=1,
Education.3=0,
Mortgage=0,
Securities.Account=0,
CD.Account=0,
Online=1,
CreditCard=1
)
```

#Normalise the new customer
```{r}
new.cust.norm<-new_customer
new.cust.norm<-predict(norm.values, new.cust.norm)
```

#Knnprediction for the new customer
```{r}
knn.pred<-class::knn(train=train.norm.df,test=new.cust.norm,cl=train.df$Personal.Loan,k=1)
knn.pred
```

2.What is a choice of k that balances between overfitting and ignoring the predictor
information?

```{r}
# Calculate the accuracy for each value of k
# Set the range of k values to consider
accuracy.df <- data.frame(k = seq(1, 15, 1), overallaccuracy = rep(0, 15))
for(i in 1:15) {
  knn.pred <- class::knn(train = train.norm.df, 
                         test = valid.norm.df, 
                         cl = train.df$Personal.Loan, k = i)
  accuracy.df[i, 2] <- confusionMatrix(knn.pred, 
                                       as.factor(valid.df$Personal.Loan),positive = "1")$overall[1]
}

which(accuracy.df[,2] == max(accuracy.df[,2])) 

plot(accuracy.df$k,accuracy.df$overallaccuracy)

```

3.Show the confusion matrix for the validation data that results from using the best k
```{r}
#considering k=3, as it has the maximum accuracy
knn.pred2<-class::knn(train=train.norm.df,test=valid.norm.df,cl=train.df$Personal.Loan,k=3)
 

confusion_matrix1 <- confusionMatrix(table(knn.pred2,valid.df$Personal.Loan))
confusion_matrix1
```
4.Consider the following customer: Age = 40, Experience = 10, Income = 84,
Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0,
Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and Credit
Card = 1. Classify the customer using the best k.


```{r}
new_customer1<- data.frame(
Age=40,
Experience=10,
Income=84,
Family=2,
CCAvg=2,
Education.1=0,
Education.2=1,
Education.3=0,
Mortgage=0,
Securities.Account=0,
CD.Account=0,
Online=1,
CreditCard=1
)

#normalize the new customer 1
new.cust.norm1<-new_customer1
new.cust.norm1<-predict(norm.values, new.cust.norm1)

#knn prediction new customer1
knn.pred3<-class::knn(train=train.norm.df,test=new.cust.norm1,cl=train.df$Personal.Loan,k=3)
knn.pred3
```
5.Repartition the data, this time into training, validation, and test sets (50% : 30% : 20%). Apply
the k-NN method with the k chosen above. Compare the confusion matrix of the test set
with that of the training and validation sets. Comment on the differences and their reason.

#split the data in to training(50%), validation(30%) and test(20%)
```{r}
set.seed(1)
```


```{r}
train.index1<-sample(row.names(universal_m.df),0.5*dim(universal_m.df)[1])
a=as.numeric(train.index1)
train.index1
a

```


```{r}
train.df1<-universal_m.df[a, ]
train.df1

tempdata<-universal_m.df[-a, ] 
tempdata
#where temp data is the remaining data to be considered for valid and test data sets.
```

#Normalise the train data
```{r}
train.norm.df <- train.df1[, -10]
norm.values1<-preProcess(train.df1[, -10],method=c("center","scale"))
train.norm.df<-predict(norm.values,train.df1[,-10])

```

#Splitting the temp data into valid and test data(20% by giving set difference of temp data from validation index)
```{r}
valid.index1<-sample(row.names(tempdata),0.6*dim(tempdata)[1])
valid.index1
test.index1 <- setdiff(row.names(tempdata), valid.index1)
test.index1

valid.df1 <- tempdata[valid.index1, ]
test.df1 <- tempdata[test.index1, ]
```


#normalize the valid and test data
```{r}
Valid.norm.df <- valid.df1[, -10]
test.norm.df <- test.df1[, -10]
norm.values2 <-preProcess(valid.df1[, -10],method=c("center","scale"))
norm.values3 <-preProcess(test.df1[, -10],method=c("center","scale"))
Valid.norm.df<-predict(norm.values,valid.df1[,-10])
test.norm.df<-predict(norm.values,test.df1[,-10])
```

#knn prediction Train, valid and test data
```{r}
knn.pred4<-class::knn(train=train.df1,test=train.df1,cl=train.df1$Personal.Loan,k=3)
knn.pred4
knn.pred5<-class::knn(train=train.df1,test=valid.df1,cl=train.df1$Personal.Loan,k=3)
knn.pred5
knn.pred6<-class::knn(train=train.df1,test=test.df1,cl=train.df1$Personal.Loan,k=3)
knn.pred6
```
#Confusion Matrix for Train, Valid and Test data

```{r}
confusion_matrix2 <- confusionMatrix(table(knn.pred4,train.df1$Personal.Loan))
confusion_matrix2
confusion_matrix3 <- confusionMatrix(table(knn.pred5,valid.df1$Personal.Loan))
confusion_matrix3
confusion_matrix4 <- confusionMatrix(table(knn.pred6,test.df1$Personal.Loan))
confusion_matrix4
```

# Compare and comment on the confusion matrices
```{r}
print("Confusion Matrix for Training set:")
print(confusion_matrix2)

print("\nConfusion Matrix for Validation Set:")
print(confusion_matrix3)

print("\nConfusion Matrix for Test Set:")
print(confusion_matrix4)
```

#Comment on the differences of training and validation sets and their reason

Below are the differences we can interpret from the above working.
Test vs Train:
Accuracy: From the above working, we can see that accuracy(0.9572) of Training confusion matrix, is slightly higher than Test Confusion matrix (0.889)

Sensitivity(True Positive Rate):Training confusion matrix has a higher sensitivity(0.9890) compared to test(0.9617)

Specificity(True Negative Rate):Both matrices have different specificity values.Training confusion matrix has a higher specificity(0.6466) compared to test confusion matrix.(0.3125)

Precision:The precision in Training confusion matrix is higher(0.9647) then test confusion matrix(0.9173).

Test vs validation:
Accuracy: From the above working, we can see that accuracy(0.9167) of validation confusion matrix, is slightly higher than Test Confusion matrix (0.889)

Sensitivity(True Positive Rate):Both matrices have almost similar sensitivity values and they both are high.Validation confusion matrix has a higher sensitivity(0.9655) compared to test(0.9617)

Specificity(True Negative Rate):Validation confusion matrix has a higher sensitivity(0.4265) compared to test(0.3125)

Precision:The precision in validation confusion matrix is higher(0.9441) then test confusion matrix(0.9173).

Therefore, on comparing we understood that training data set has the highest accuracy when compared to test and validation data sets, which indicates that algorithm is operating as intended.
