---
title: "Clustering"
author: '"Pallavi"'
date: "2023-11-12"
output: html_document
---

#Running the required libraries
```{r}
library(ggplot2)
library(factoextra)
library(flexclust)
library(cluster)
library(tidyverse)
```

#Import and read the CSV file
```{r}
pharmaceuticalindustry <- read.csv("C:\\Users\\palla\\OneDrive\\Desktop\\Assignments\\FML\\Last assignment\\Pharmaceuticals.csv")
dim(pharmaceuticalindustry)
summary(pharmaceuticalindustry)
```

#   1.Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) used, the number of clusters formed, and so on. 

```{r}
pharmaceuticalindustry <- na.omit(pharmaceuticalindustry) #provides us with the data post eliminating the incomplete cases.
pharmaceuticalindustry

#using only the quantitative variables(1-9) to cluster the 21 firms
row.names(pharmaceuticalindustry)<- pharmaceuticalindustry [,1]
pharmaceuticalindustry1<- pharmaceuticalindustry [,3:11]
head(pharmaceuticalindustry1)

#Now, Scale all the quantitative variables in the dataframe
pharmaceuticalindustry2<-scale(pharmaceuticalindustry1)
head(pharmaceuticalindustry2)

#Further more,to determine the no of clusters to do the cluster analysis using Elbow Method
fviz_nbclust(pharmaceuticalindustry2, kmeans, method = "wss") + labs(subtitle = "Elbow Method")

#Silhouette method for determining no of clusters
fviz_nbclust(pharmaceuticalindustry2, kmeans, method = "silhouette")+ labs(subtitle = "Silhouette Method")

fviz_nbclust(pharmaceuticalindustry2, kmeans, method = "gap_stat") + labs(subtitle = "Gap Stat Method")
```
Therefore, from the above plots, it is clear that the no of clusters are 5 and they are enough to show the variations that are present in the data


```{r}
set.seed(120)
k5 <- kmeans(pharmaceuticalindustry2, centers = 5, nstart = 25)
#Visualize the output
k5$centers  #centroids
fviz_cluster(k5,data = pharmaceuticalindustry2) # to Visualize the clusters
k5
distance<- dist(pharmaceuticalindustry2, method = "euclidean")
fviz_dist(distance)

#K-Means Cluster Analysis- Fit the data with 5 clusters
Fitting<-kmeans(pharmaceuticalindustry2,5)

#Finding the mean value of all quantitative variables for each cluster
aggregate(pharmaceuticalindustry2,by=list(Fitting$cluster),FUN=mean)

pharmaceuticalindustry3<-data.frame(pharmaceuticalindustry2,Fitting$cluster)
pharmaceuticalindustry3

#To view the cluster plot
library(cluster)
clusplot(pharmaceuticalindustry2,Fitting$cluster, color = TRUE, shade = TRUE, 
         labels = 2,
         lines = 0)
```

#2.Interpret the clusters with respect to the numerical variables used in forming the clusters. Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used in forming the clusters)



```{r}
aggregate(pharmaceuticalindustry2, by = list(Fitting$cluster), FUN = mean)
Pharmacy <- data.frame(pharmaceuticalindustry2,k5$cluster)
Pharmacy
```
Cluster 1 - BAY, CHTT, IVX
Cluster 1 has highest Beta , Leverage and lowest Market_Cap, ROE, ROA, Leverage, Rev_Growth, Net_Profit_Margin

Cluster 2 - ABT, AZN, BMY, LLY, NVS, SGP,WYE
Cluster 2 has highest Net_Profit_Margin and lowest Beta.

Cluster 3 - ELN, MRX
Cluster 3 has highest Rev_Growth and lowest PE_Ratio, Asset_Turnover.

Cluster 4 - JNJ, MRK, PFE, GSK
Cluster 4 has highest Market_Cap, ROE, ROA,Asset_Turnover 

Cluster 5 - AGN, AHM, AVE, PHA, WPI
Cluster 5 has highest PE_Ratio.

#Is there a pattern in the clusters with respect to the numerical 
#variables (10 to 12)? (those \n #not used in forming the clusters)

```{r}
Pharma <- pharmaceuticalindustry[12:14] %>% mutate(Clusters=k5$cluster)
ggplot(Pharma, mapping = aes(factor(Clusters), fill =Median_Recommendation))+geom_bar(position='dodge')+labs(x ='Clusters',y ='Frequence')
ggplot(Pharma, mapping = aes(factor(Clusters),fill = Location))+
  geom_bar(position = 'dodge')+labs(x ='Clusters',y = 'Frequence')
ggplot(Pharma, mapping = aes(factor(Clusters),fill = Exchange))+geom_bar(position = 'dodge')+
  labs(x ='Clusters',y = 'Frequence')
```
As can be seen from the graphs above, each cluster has a narrow pattern.

#The companies in cluster 1 are evenly distributed throughout AMEX, NASDAQ, and NYSE, despite having a different Hold and Moderate Buy median, a different count from the US and Germany, and a distinct nation count.

#The cluster 2 has equal Hold and Moderate Buy medians, is available solely on the NYSE, and is distributed equally throughout the US and Canada.

#In addition to equal Moderate Buy and Sell medians, Cluster 3 is a stock that is listed on the NYSE and has unique counts for France, Ireland, and the US.

#In Cluster 4, the medians for Hold, Strong Buy, and Moderate Buy are arranged from highest to lowest. Originating in the US, the UK, and Switzerland, they are listed on the NYSE.

#In Cluster 5's distribution across the US and the UK, as well as its NYSE listing, it shares similar hold and mild buy medians.


Could see a pattern among the clusters with respect to variables(10 to 12)

Clusters 3,4,5 has mostly Moderate Buy Recommendation

Clusters 1,2 has Hold Recommendation


#3.Provide an appropriate name for each cluster using any or all of the variables in the dataset.
Cluster1 - high Beta, Leverage cluster (or) Buy Cluster. 

Cluster2 - high Net_Profit_Margin cluster (or) high hold cluster.

Cluster3 - Low PE_Ratio, Asset_Turnover cluster (or) hold cluster.

Cluster4 - Moderate Buy cluster

Cluster5 - high PE_Ratio cluster (or) high Buy cluster.

